Lesson 3
========================================================

***

### Pseudo-Facebook User Data

```{r Pseudo-Facebook User Data}
pf <- read.csv('pseudo_facebook.tsv', sep ='\t')
names(pf)
```

***

### Histogram of Users' Birthdays

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
qplot(x=dob_day, data=pf) + scale_x_discrete(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:
There are many more people born on the 1st of the month. I would expect that the number of people born on each day of the month to be about the same, except the 31st since only a few months have 31 days.

***
### Faceting
```{r Faceting}
qplot(x=dob_day, data=pf, binwidth=1) + facet_wrap(~dob_month, ncol=4)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:
There are many many more people born on January 1 than any other day of the year.
***

### Be Skeptical - Outliers and Anomalies
Notes:
The large numbers on January 1 could be an error or users defaulting to January 1 for privacy.

***

### Friend Count
#### What code would you enter to create a histogram of friend counts?
```{r Friend Count}
qplot(x=friend_count, data = pf)
```

### Limiting the Axes
```{r Limiting the Axes}
qplot(x=friend_count, data = pf) + scale_x_continuous(limits = c(0,1000))
```

### Faceting Friend Count
####What code would you add to create a facet the histogram by gender?
```{r Faceting Friend Count}
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + facet_wrap(~gender)
```

***

### Omitting NA Values
```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + facet_wrap(~gender)
```

***

### Statistics 'by' Gender
```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:
Women
#### What's the difference between the median friend count for women and men?
Response:
22
#### Why would the median be a better measure than the mean?
Response:
There are some people with many more or far fewer friends than the majority of the group, which can skew the mean if they are not balanced. The median is more robust to individual responses.
***

### Tenure
```{r Tenure}
qplot(x=tenure, data=pf, binwidth = 30, color = I('black'), fill = I('#099DD9'))
```

***

#### How would you create a histogram of tenure by year?
```{r Tenure Histogram by Year}
qplot(x=tenure/365, data=pf, binwidth = .25, color = I('black'), fill = I('#099DD9'))
```

***

### Labeling Plots
```{r Labeling Plots}
qplot(x=tenure/365, data=pf, binwidth = .25, color = I('black'), fill = I('#F79420')) + xlab('Number of years using Facebook') + ylab('Number of users in sample') + scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0,7))
```

***

### User Ages
```{r User Ages}
qplot(x=age, data=pf, binwidth = 1, color = I('black'), fill = I('#5760AB')) + xlab('Ages of users in sample') + ylab('Number of users in sample') + scale_x_continuous(breaks = seq(0, 120, 20), limits = c(10,125))
```

#### What do you notice?
Response:
There are some spikes in ages > 100
***

### Transforming Data
```{r}
#install.packages("gridExtra")
library(gridExtra) 
fcount <- qplot(x=friend_count, data = pf) + geom_histogram()
fcountlog <- fcount + scale_x_log10()
fcountsqrt<- fcount + scale_x_sqrt()
grid.arrange(fcount, fcountlog, fcountsqrt, ncol=1)
```

***

### Add a Scaling Layer
```{r Add a Scaling Layer}
fcountlog <- qplot(x=log10(friend_count + 1), data = pf)
fcountlog2 <- qplot(x=friend_count, data = pf) + scale_x_log10()
grid.arrange(fcountlog, fcountlog2, ncol=2)
```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x=friend_count, data=subset(pf, !is.na(gender)), bindwidth=10, geom='freqpoly', color=gender) + scale_x_continuous(lim=c(0,1000), breaks=seq(0,1000,50))

qplot(x=friend_count, y=..count../sum(..count..), data=subset(pf, !is.na(gender)), bindwidth=10, geom='freqpoly', color=gender) + scale_x_continuous(lim=c(0,1000), breaks=seq(0,1000,50)) + xlab('Friend Count') + ylab('Proportion of Users With That Friend Count')
```

***

### Likes on the Web
```{r Likes on the Web}
qplot(x=www_likes, data=subset(pf, !is.na(gender)), geom='freqpoly', color=gender) + scale_x_continuous() + scale_x_log10()
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
```{r Box Plots}
qplot(x=gender, y=friend_count, data=subset(pf, !is.na(gender)), geom='boxplot')
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.
```{r}
qplot(x=gender, y=friend_count, data=subset(pf, !is.na(gender)), geom='boxplot') + coord_cartesian(ylim = c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
```{r Box Plots, Quartiles, and Friendships}
qplot(x=gender, y=friend_count, data=subset(pf, !is.na(gender)), geom='boxplot') + coord_cartesian(ylim = c(0,250))
by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:women
#### Write about some ways that you can verify your answer.
Response:
We can find the quantiles by summary and boxplots.
```{r Friend Requests by Gender}
by(pf$friendships_initiated, pf$gender, summary)
qplot(x=gender, y=friendships_initiated, data=subset(pf, !is.na(gender)), geom='boxplot') + coord_cartesian(ylim = c(10,125))
```

***

### Getting Logical
```{r Getting Logical}
summary(pf$mobile_likes)
pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes>0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

length(subset(pf$mobile_check_in, pf$mobile_check_in == 1))/length(pf$mobile_check_in)*100.
```