---
title: "Problem Set 3"
output: html_document
---

## Load the diamonds data set
```{r}
library(ggplot2)
data(diamonds)
summary(diamonds)
dim(diamonds)
```
## Create a histogram of the price of all the diamonds in the diamond data set
```{r}
qplot(data=diamonds, x=price, binwidth=500)
dim(subset(diamonds, diamonds$price < 500))[1]
dim(subset(diamonds, diamonds$price < 250))[1]
dim(subset(diamonds, diamonds$price >= 15000))[1]
```

## Explore the largest peak in the price histogram you created earlier.Try limiting the x-axis, altering the bin width, and setting different breaks on the x-axis.
```{r}
qplot(data=diamonds,x=price,binwidth=50) + scale_x_continuous(limits=c(0,1500))
ggsave('priceHistogram.png')
```

## Break out the histogram of diamond prices by cut.
```{r}
qplot(data=diamonds,x=price, binwidth=500) + facet_wrap(~cut, scales="free")
```

## Create a histogram of price per carat and facet it by cut. Adjust the bin width and transform the scale of the x-axis using log10.
```{r}
qplot(data=diamonds,x=price/carat,binwidth=.02) + facet_wrap(~cut,scales = "free") + scale_x_log10()
```

## Investigate the price of diamonds using box plots, numerical summaries, and one of the following categorical variables: cut, clarity, or color.
```{r}
a<-qplot(x=clarity,y=price, data=diamonds, geom='boxplot') + coord_cartesian(ylim = c(0,7000))
b<-qplot(x=color,y=price, data=diamonds, geom='boxplot') +
  coord_cartesian(ylim = c(0,8000))
c<-qplot(x=cut,y=price, data=diamonds, geom='boxplot') +
  coord_cartesian(ylim = c(0,7000))

library(gridExtra) 
grid.arrange(a,b,c,ncol=1)
ggsave('priceBoxPlots.png')
```

```{r}
by(diamonds$price,diamonds$color,summary)
IQR(subset(diamonds, color=='D')$price)
IQR(subset(diamonds, color=='J')$price)
```

## Investigate the price per carat of diamonds across the different colors of diamonds using boxplots.
```{r}
qplot(x=color,y=price/carat, data=diamonds, geom='boxplot') + coord_cartesian(ylim = c(2000,6000))
```
```{r}
qplot(x=carat, data=diamonds, geom='freqpoly', binwidth = .5)
qplot(x=carat, data=diamonds, geom='freqpoly', binwidth = .1) + scale_x_continuous(limits = c(0, 5), breaks = seq(0, 5, .1))
```

##Your task is to investigate the distribution of your friends' birth months and days. Here some questions you could answer:

###How many people share your birthday?
### Which month contains the most number of birthdays?
### How many birthdays are in each month?
### Which day of the year has the most number of birthdays?

```{r}
df <-read.csv('birthdaysExample.csv')
df$month <- strftime(strptime(df$dates, format="%m/%d/%y"),"%m") 
df$day <- strftime(strptime(df$dates, format="%m/%d/%y"),"%d") 

dim(subset(df, df$month=="01" & df$day =="01"))[1]

library(dplyr)
df.fc_by_birthmonth <- df %>%
   group_by(month) %>%
   summarise(n=n()) %>%
  arrange(month)

subset(df.fc_by_birthmonth, df.fc_by_birthmonth$n==max(df.fc_by_birthmonth$n))$month
df.fc_by_birthmonth

df.fc_by_birthday <- df %>%
   group_by(month, day) %>%
   summarise(n=n()) %>%
  arrange(month, day)
subset(df.fc_by_birthday, df.fc_by_birthday$n==max(df.fc_by_birthday$n))
```
