Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
```{r Scatterplots}
library(ggplot2)
df <- read.csv('pseudo_facebook.tsv', sep='\t')
qplot(x=age, y=friend_count, data=df)
```

***

#### What are some things that you notice right away?
Response:
There seems to a general larger spread amongst younger users.
***

### ggplot Syntax
```{r ggplot Syntax}
ggplot(aes(age,friend_count), data=df) + geom_point() + xlim(13,90)
summary(df$age)
```

***

### Overplotting
```{r Overplotting}
ggplot(aes(age,friend_count), data=df) + geom_jitter(alpha=.05) + xlim(13,90)
```

***

### Coord_trans()
#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(age,friend_count), data=df) + geom_point(alpha=.05, position=position_jitter(h=0)) + xlim(13,90) + coord_trans(ytrans='sqrt')
```

#### What do you notice?
The shape of the data changed. The older you are, the less likely you are to have few friends.
***

### Alpha and Jitter
```{r Alpha and Jitter}
ggplot(aes(age,friendships_initiated), data=df) + geom_point(alpha=.05, position=position_jitter(h=0)) + xlim(13,90) + coord_trans(ytrans='sqrt')
```

***

### Overplotting and Domain Knowledge

***

### Conditional Means
```{r Conditional Means}
library(dplyr)
age_groups <- group_by(df, age)
df.fc_by_age <- df %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n=n()) %>%
            arrange(age)
head(df.fc_by_age)
```

Create your plot!
```{r Conditional Means Plot}
ggplot(aes(age,friend_count_mean), data=df.fc_by_age) + geom_line()
```

***

### Overlaying Summaries with Raw Data
```{r Overlaying Summaries with Raw Data}
ggplot(aes(age,friend_count), data=df) + geom_point(alpha=0.05, position=position_jitter(h=0)) + geom_point(alpha=0.05, position = position_jitter(h=0), color='orange') + geom_line(stat='summary', fun.y=mean) + geom_line(stat='summary', fun.y=quantile, probs=.1, linetype = 2, color = 'blue') + geom_line(stat='summary', fun.y=quantile, probs=.9, linetype = 2, color = 'blue') + geom_line(stat='summary', fun.y=quantile, probs=.5, color = 'blue') + coord_cartesian(xlim = c(13, 70), ylim=c(0,1000))
```

#### What are some of your observations of the plot?
Response:
The median friend count doesn't change much with age.
***

### Correlation
```{r Correlation}
cor.test(df$age, df$friend_count, method='pearson')
with(df, cor.test(age,friend_count, method='pearson'))
```

What's the correlation between age and friend count? Round to three decimal places.
Response:
-0.027 
***

### Correlation on Subsets
```{r Correlation on Subsets}
with(subset(df, age<=70), cor.test(age, friend_count))
```

***

## Create Scatterplots
```{r}
ggplot(aes(www_likes_received, likes_received), data=df) + geom_point()
```

***

### Strong Correlations
```{r Strong Correlations}
ggplot(aes(www_likes_received, likes_received), data=df) + geom_point() + xlim(0, quantile(df$www_likes_received, .95)) + ylim(0, quantile(df$www_likes, .95)) + geom_smooth(method='lm',color='red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(df$www_likes, df$www_likes_received, method ='pearson')
```

***

### More Caution with Correlation
```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
```

Create your plot!
```{r Temp vs Month}
data("Mitchell")
ggplot(aes(Month, Temp), data=Mitchell) + geom_point()
```

***

### Noisy Scatterplots 

a. Take a guess for the correlation coefficient for the scatterplot.
0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
```{r Making Sense of Data}
ggplot(aes(Month, Temp), data=Mitchell) + geom_point() + scale_x_discrete(breaks = seq(0,203,12)) + coord_fixed(ratio=.5)
```

***

### A New Perspective
What do you notice?
Response:
There's a cyclical pattern!

***

### Understanding Noise: Age to Age Months
```{r Understanding Noise: Age to Age Months}
df$age_with_months <- df$age + (12 - df$dob_month)/12.
```

***

### Age with Months Means
```{r Age with Months Means}
df.fc_by_age_months <- df %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n=n()) %>%
            arrange(age_with_months)
```

***

### Noise in Conditional Means
```{r Noise in Conditional Means}
ggplot(aes(age_with_months,friend_count_mean), data = subset(df.fc_by_age_months, age_with_months<71)) + geom_point()
```

***

### Smoothing Conditional Means
```{r Smoothing Conditional Means}
ggplot(aes(age_with_months,friend_count_mean), data = subset(df.fc_by_age_months, age_with_months<71)) + geom_point() + geom_smooth()
```