---
title: "Problem Set 4"
output: html_document
---

####Your first task is to create a scatterplot of price vs x.
```{r}
library(ggplot2)
data(diamonds)
summary(diamonds)
ggplot(data=diamonds, aes(price,x)) + geom_point()
```

```{r}
cor.test(diamonds$price, diamonds$x)
cor.test(diamonds$price, diamonds$y)
cor.test(diamonds$price, diamonds$z)
```
#### Create a simple scatter plot of price vs depth
```{r}
ggplot(data=diamonds, aes(depth, price)) + geom_point(alpha=0.01) + scale_x_continuous(breaks = seq(40,80,2))
cor.test(diamonds$depth,diamonds$price)
```
#### Create a scatterplot of price vs carat and omit the top 1% of price and carat values.
```{r}
ggplot(data=diamonds, aes(carat, price)) + geom_point(alpha=0.01) + xlim(0, quantile(diamonds$carat, .99)) + ylim(0, quantile(diamonds$price, .99)) 
```

#### Create a scatterplot of price vs. volume (x * y * z). This is a very rough approximation for a diamond's volume.
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z
ggplot(data=diamonds, aes(volume, price)) + geom_point()
library(plyr)
count(diamonds$volume == 0)
```
#### Subset the data to exclude diamonds with a volume greater than or equal to 800. Also, exclude diamonds with a volume of 0. Adjust the transparency of the points and add a linear model to the plot.
```{r}
with(subset(diamonds, volume!=0 & volume < 800), cor.test(price,volume))

ggplot(data=subset(diamonds, volume!=0 & volume < 800), aes(volume, price)) + geom_point(alpha=0.01) + geom_smooth(method='lm',color='red')
```

#### Use the function dplyr package to create a new data frame containing info on diamonds by clarity. Name the data frame diamondsByClarity

The data frame should contain the following variables in this order.

       (1) mean_price
       (2) median_price
       (3) min_price
       (4) max_price
       (5) n
where n is the number of diamonds in each level of clarity.

```{r}
detach("package:plyr", TRUE)
library(dplyr)
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price)*1.,
            min_price = min(price),
            max_price = max(price),
            n = n()) %>%
  arrange(clarity)
```

#### We’ve created summary data frames with the mean price by clarity and color. You can run the code in R to  verify what data is in the variables diamonds_mp_by_clarity and diamonds_mp_by_color.

Your task is to write additional code to create two bar plots on one output image using the grid.arrange() function from the package gridExtra.

```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))
p1 <- qplot(clarity, data=diamonds_mp_by_clarity, geom ="bar", weight=mean_price)

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
p2 <- qplot(color, data=diamonds_mp_by_color, geom ="bar", weight=mean_price)

diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))
p3 <- qplot(cut, data=diamonds_mp_by_cut, geom ="bar", weight=mean_price)

library(gridExtra)
grid.arrange(p1,p2,p3, ncol=1)
```